Impact offsetting

### 2022-02-14 paper.tex
Two separate problems perspective

\section{Stochastic side-effect avoidance}
	This section presents an as of yet unexplored choice for $\baseline$.
	\subsection{Basic idea}
	 We would like to answer the counterfactual question ``What would the world look like if we used
		the baseline policy rather than the policy we actually used?''.
	 A solution to this problem would be a generator of baseline states that would,
		given an arbitrary state $s$, produce a distribution of states that could have
		been encountered instead of $s$.

@@

		The resulting generator could then be used as $\baseline$ with an arbitrary side-effect avoidance
		measure, like relative reachability. For a given state it should be likely to return
		a baseline state corresponding to a possible rollout of the baseline policy from possibly some
		steps in the past.

		\subsubsection{Doubts}
		 Does this actually work?
			\begin{enumerate}
				\item Does training with $L$ actually create a generator that
					generates things mostly corresponding to the $l$th argument with about $2^{-l}$
					probability? Or does this converge to something else, e.g. always the $0$th argument?
				\item Is it reasonable expect that the CVAE model will be able to ``model'' the whole
					transformation needed to get from $s_t$ to $s^{\left( t-4^l \right)}_{4^l}$?
					After all this requires going back $4^l$ steps and back forward using a different policy.
					Even if we had one model per $l$ this might be a nontrivial task. On the other hand this might
					be simpler if the baseline policy and undisturbed environment behaviour is simple.
				\item Variant of the above -- if we use multiple models we could have them generate
					partial (as in only some steps) trajectories. This I at least know is mostly possible
					and might improve accuracy and interpretability.
				\item What about parts of the state space that random policies cannot reach, but the
					actual policy used by the training agent can. How well will this generalize?
			\end{enumerate}
\appendix
	\section{Non-Markovian rewards}
	 In the main text I pretended everything is Markovian, but at least Krakovna's original approach
		is not. A definition that would actually apply to her appraoch is presented below,
		but was too unwieldy to work with and did not actually provide much insight.
		\begin{definition}[Side effect avoidance]
			\emph{Side effect avoidance} is a class of approaches for mitigating reward misspecification, which
			modify the reward passed to the agent according to the equation
			\begin{equation}
				\modifiedRewardFunction\left( \left( s_i \right)_{i=0}^t \right) =
				\corruption\left( s_t \right) -
				\lambda \cdot \deviationFromBaseline\left( \left( s_i \right)_{i=0}^t, \baseline\left( \left( s_i \right)_{i=0}^t \right) \right),
				\label{eq:sideEffectAvoidance-nonMarkov}
			\end{equation}
			where $t$ is the time step, $s_i$ is the state at time $i$, $\baseline$ is a probability
			distribution of trajectories of the baseline agent, and $\deviationFromBaseline$ is a measure
			of the deviation between a trajectory and a baseline distribution of trajectories.

>>>>

\section{Stochastic side effect avoidance}
	This section presents an as of yet unexplored choice for $\baseline$.
	\subsection{Basic idea}
	 We would like to answer the counterfactual question ``What would the world look like if we used
		the baseline policy rather than the policy we actually used?''.
	 A solution to this problem would be a generator of baseline states that would,
		given an arbitrary state $s$, produce a distribution of states that could have
		been encountered instead of $s$.

@@

		The resulting generator could then be used as $\baseline$ with an arbitrary side effect avoidance
		measure, like relative reachability. For a given state it should be likely to return
		a baseline state corresponding to a possible rollout of the baseline policy from possibly some
		steps in the past.

		\subsubsection{Doubts}
		 Does this actually work?
			\begin{enumerate}
				\item Does training with $L$ actually create a generator that
					generates things mostly corresponding to the $l$th argument with about $2^{-l}$
					probability? Or does this converge to something else, e.g. always the $0$th argument?
				\item Is it reasonable expect that the CVAE model will be able to ``model'' the whole
					transformation needed to get from $s_t$ to $s^{\left( t-4^l \right)}_{4^l}$?
					After all this requires going back $4^l$ steps and back forward using a different policy.
					Even if we had one model per $l$ this might be a nontrivial task. On the other hand this might
					be simpler if the baseline policy and undisturbed environment behaviour is simple.
				\item Variant of the above -- if we use multiple models we could have them generate
					partial (as in only some steps) trajectories. This I at least know is mostly possible
					and might improve accuracy and interpretability.
				\item What about parts of the state space that random policies cannot reach, but the
					actual policy used by the training agent can. How well will this generalize?
			\end{enumerate}
\section{Are we trying to solve two problems at once?}
	The approaches used by Krakovna and Turner, due to their choices for $\deviationFromBaseline$,
	can be summed up as ``Avoid increasing power compared to the baseline.''. This is a pretty natural
	result of trying to avoid both breaking vases (decreasing everyone's power) and stealing sushi
	(increasing the agent's power at the cost of the hungry human). However, it has the inconvenient
	consequence of making us work with nonlinear functions, something everybody hates.
	An alternative approach is to treat this problem as two separate problems.

	First problem, we want the agent to have power. If it is incentivized to seek power in general,
	for example using the future tasks paradigm (but not relative to any baseline), it will avoid
	performing irreversible actions, such as breaking vases. A natural transformation of the approaches
	of Krakovna and Turner to solving this problem ends up massively simpler than side effect avoidance.
	We just have to add a term for increasing power to the reward, using future tasks would be linear,
	and we wouldn't even have to worry about any baselines.

	Second problem, avoiding interference with other agents. Achieving any specific goal almost always
	requires sacrificing some power, like eating sushi prevents you from using it later. Agents will
	thus routinely lower everyone's power when achieving their goals. An agent incentivized to increase
	power would also be incentivized to interfere with all other agents. Preventing this is much harder
	than the solution to the previous problem, consider for example that agent $A$ cannot even rely
	on agent $B$'s presence as a clue to whether a specific power-lowering phenomenon is intended by $B$.
	This would work for eating sushi, but not for automated machinery set up in a factory before agent $A$
	existed. Considering how much simpler the solution to the first problem is than side effect avoidance,
	investigating this one might still be worthwhile.

	Note that the approach intensifies a problem in AI alignment commonly mentioned since the very
	inception of the field -- power seeking is an instrumental goal for sufficiently general
	agents anyway, so it seems unwise to additionally encourage it.

	On the other hand this might actually increase the capabilities of agents, by explicitly introducing
	incentives we expect in more capable agents.
	Has anyone checked that? Worrying thought.
\appendix
	\section{Non-Markovian rewards}
	 In the main text I pretended everything is Markovian, but at least Krakovna's original approach
		is not. A definition that would actually apply to her appraoch is presented below,
		but was too unwieldy to work with and did not actually provide much insight.
		\begin{definition}[Side effect avoidance]
			\emph{Side effect avoidance} is a class of approaches for mitigating reward misspecification, which
			modify the reward passed to the agent according to the equation
			\begin{equation}
				\modifiedRewardFunction\left( \left( s_i \right)_{i=0}^t \right) =
				\corruption\left( s_t \right) -
				\lambda \cdot \deviationFromBaseline\left( \left( s_i \right)_{i=0}^t, \baseline\left( \left( s_i \right)_{i=0}^t \right) \right),
				\label{eq:sideEffectAvoidance-nonMarkov}
			\end{equation}
			where $t$ is the time step, $s_i$ is the state at time $i$, $\baseline$ is a probability
			distribution of trajectories of the baseline agent, and $\deviationFromBaseline$ is a measure
			of the deviation between a trajectory and a baseline distribution of trajectories.
